Data preparation

### 2020-07-28 docs/paper.Rmd
update-double-render

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use,  empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy. Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::html_document2
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
# knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
pacman::p_load(knitr,kableExtra, dplyr, captioner)
```

@@

```{r include=FALSE}
table_nums <- captioner(prefix = "Table")
fig_nums   <- captioner(prefix = "Figure")
anx_nums<-captioner(prefix = "Annexed")
```

@@

# Introducción

@@

## The black-box of meritocratic beliefs

@@

## An instrument proposal

@@

```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}

knitr::include_graphics('../input/images/generalf.png')
```

@@

## Data collection

@@

The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014]. The eight items ordered according to dimensions are presented in Table 2. These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).

```{r include=FALSE}
cap <- "Items according to dimension."
if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
```

```{r echo=FALSE}
 tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",")

 cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")

 kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 14, bootstrap_options=c("striped", "bordered")) %>%
   column_spec(column = 1,width = "1.5cm",) %>%
   column_spec(column = 2,width = "2 cm") %>%
   collapse_rows(columns = 1:2,valign = "middle")

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in __Figure 1__. The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

```{r include=FALSE}
cap <- "Survey flow"
if (knitr::is_html_output()) {diam1 <- fig_nums("diam1",cap)} else {diam1 <- cap}
```

![`r diam1`](../input/images/app_mod.png){ width=80% }

## Methods

@@

```{r include=FALSE}
cap <- "Theorical model"
if (knitr::is_html_output()) {diam2 <- fig_nums("diam2",cap)} else {diam2 <- cap}
```

![`r diam2`](../output/images/meas01.png)

@@

Como puede observarse en la tabla 3, los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo). Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.

```{r include=FALSE}
cap <- "Descriptive statistics of the scale."
if (knitr::is_html_output()) {tades <- table_nums("tades",cap)} else {tades <- cap}
```

```{r echo=FALSE}
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
                          "B.Perception Talent" = perc_talent,
                          "C.Perception rich parents" = perc_wpart,
                         "D.Perception contacts" = perc_netw,
                          "E.Preferences Effort" = pref_effort,
                          "F.Preferences Talent" =  pref_talent,
                         "G.Preferences rich parents" =  pref_wpart ,
                          "H.Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
) %>% kable_styling(full_width = table_format2, font_size = 14,position = "center");sumdat04
```

El gráfico XX muestra ...
A partir del gráfico se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.

```{r include=FALSE}
cap <- "Descriptive plot"
if (knitr::is_html_output()) {liker <- fig_nums("liker",cap)} else {liker <- cap}
```

<center>

![`r liker`](../output/images/plotlikert.png){ width=60% }

</center>

El Grafico XX muestra ....
Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces. Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

```{r include=FALSE}
cap <- "polychoric correlation plot"
if (knitr::is_html_output()) {fb <- fig_nums("fb",cap)} else {fb <- cap}
```

<center>

![`r fb`](../output/images/corpoly.png){ width=60% }

</center>

@@

Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla XX

```{r include=FALSE}
cap <- "Factor loads and model fit."
if (knitr::is_html_output()) {loadfit <- table_nums("loadfit",cap)} else {loadfit <- cap}
```

```{r echo=FALSE}

load(file = "../output/tables/tb.fm.RData")

tb.col  <- c("Variables",1:4,1:4,1:4)
tb.caption <- "Factor loadings and fit measures"
tb.foot <- paste0("Standardised factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")

@@

tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = loadfit) %>%
          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
          row_spec(row = 8,hline_after = TRUE) %>%
          add_indent(c(9:12))  %>%
          footnote(general = tb.foot,threeparttable = T);tb.fm01

@@

```{r include=FALSE}
cap <- "Items according to dimension."
if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
```

|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
|:---------:|:-----------------------:|:---------------------------:|:---------------:|
| **n**     |           1769          |             1769            |       1769      |
| **CFI**   |          0.595          |            0.988            |      0.994      |
| **TLI**   |          0.433          |            0.976            |      0.985      |
| **RMSEA** |          0.226          |            0.047            |      0.036      |
| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
| **p**     |           .000          |             .000            |       .000      |
| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
Table:`r te`

@@

# Anexos

```{r include=FALSE}
cap <- "Representativeness of the sample."
if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
```

|                     | Sample  |   CEP  |
|---------------------|:-------:|:------:|
| __gender__          |         |        |
| Men                 | 49,82%  | 50,52% |
| Women               | 50.18%  | 49,47% |
| __Age__             |         |        |
| 18 - 24             | 18,55%  | 18,17% |
| 25 - 34             | 18,86%  | 17,48% |
| 35 - 44             | 19.09%  | 19,98% |
| 45 - 54             | 17,96%  | 19,23% |
| 55 - or more        | 25,54%  | 25.11% |
| __Education__       |         |        |
| Primary  or less    | 2,93%   | 15,88% |
| Hig school          | 43,23%  | 37,04% |
| Non university      | 32,63%  | 28,93% |
| university or more  | 21,21%  | 18,13% |
Table:`r aax`

# Bibliografía

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use,  empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy. Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false 
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '75%',fig.pos= "H")
# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
pacman::p_load(knitr,kableExtra, dplyr)
```

@@

@@

# (PART\*) Introducción {-}

@@

## The black-box of meritocratic beliefs {-#blackbox}

@@

## An instrument proposal {-#instrumentprop}

@@

```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
knitr::include_graphics("../input/images/generalf.png")
```

@@

## Data collection {-#data}

@@

The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014]. The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores). These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).

```{r include=FALSE}
cap <- "Items according to dimension."
# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
```

```{r table-indicadores, echo=FALSE}
tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")

kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kable_styling(
    full_width = T,
    latex_options = c("hold_position", "scale_down"),
    position = "center",
    font_size = 10,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", ) %>%
  column_spec(column = 2,width = "2 cm") %>%
  collapse_rows(columns = 1:2,valign = "middle")
```

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod) . The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
knitr::include_graphics('../input/images/app_mod.png')
```

# Methods

@@

```{r meas01, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
knitr::include_graphics('../output/images/meas01.png')
```

@@

Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo). Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
                          "B.Perception Talent" = perc_talent,
                          "C.Perception rich parents" = perc_wpart,
                         "D.Perception contacts" = perc_netw,
                          "E.Preferences Effort" = pref_effort,
                          "F.Preferences Talent" =  pref_talent,
                         "G.Preferences rich parents" =  pref_wpart ,
                          "H.Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
```

El gráfico XX muestra ...
A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.

```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
knitr::include_graphics('../output/images/plotlikert.png')
```

El Grafico \@ref(fig:corpoly) muestra ....
Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces. Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
knitr::include_graphics('../output/images/corpoly.png')
```

@@

> Julio: Esta sección debe ser revisada:

- CFA general
- Fit x modelo

[Aquí SEM plot con cargas y ajuste global del modelo]

Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \@ref(tab:tab-loads)

```{r tab-loads, echo=FALSE}
load(file = "../output/tables/tb.fm.RData")
cap <- "Factor loads and model fit."
tb.col  <- c("Variables",1:4,1:4,1:4)
tb.caption <- "Factor loadings and fit measures"
tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")

@@

tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
          row_spec(row = 8,hline_after = TRUE) %>%
          add_indent(c(9:12))  %>%
          footnote(general = tb.foot,threeparttable = T);tb.fm01
```

@@

> Julio: ¿Esto se debe mantener?

```{r include=FALSE}
cap <- "Items according to dimension."
# if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
```

|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
|:---------:|:-----------------------:|:---------------------------:|:---------------:|
| **n**     |           1769          |             1769            |       1769      |
| **CFI**   |          0.595          |            0.988            |      0.994      |
| **TLI**   |          0.433          |            0.976            |      0.985      |
| **RMSEA** |          0.226          |            0.047            |      0.036      |
| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
| **p**     |           .000          |             .000            |       .000      |
| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
Table:`r cap`

@@

# References

<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->

# Appendix {-}

```{r include=FALSE}
cap <- "Representativeness of the sample."
# if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
```

|                     | Sample  |   CEP  |
|---------------------|:-------:|:------:|
| __Gender__          |         |        |
| Men                 | 49,82%  | 50,52% |
| Women               | 50.18%  | 49,47% |
| __Age__             |         |        |
| 18 - 24             | 18,55%  | 18,17% |
| 25 - 34             | 18,86%  | 17,48% |
| 35 - 44             | 19.09%  | 19,98% |
| 45 - 54             | 17,96%  | 19,23% |
| 55 - or more        | 25,54%  | 25.11% |
| __Education__       |         |        |
| Primary  or less    | 2,93%   | 15,88% |
| Hig school          | 43,23%  | 37,04% |
| Non university      | 32,63%  | 28,93% |
| university or more  | 21,21%  | 18,13% |
Table:`r cap`

### 2020-07-28 production/prod_analysis-cfa.rmd
update-double-render

---
title: "Analysis PPM-S"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

>>>>

---
title: "Analysis PPM-S"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

### 2020-07-28 production/prod_prep-data.rmd
update-double-render

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/bib/style.css"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

>>>>

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/bib/style.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---
